Student Log-Data from a Randomized Evaluation of Educational Technology: A Causal Case Study

### 2020-01-07 paper.tex
updated treatment effect description
comments 37-end, new conclusion, spellcheck

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

@@

Table \ref{tab:matchResults} gives four treatment effect estimates
using these methods.
The first row gives the result using precision-weights without
regression adjustment. estimating an effect of $H$
The rows labeled ATE and TOT weight the matched sets to
estimate the average effect of $H$ over the entire treatment
group and over those members of the treatment group with $H=1$,
respectively.
The row labeled ANCOVA of Table \ref{tab:matchResults} gives the
coefficient on $H$ from a regression of Y on $H$ along with indicators for matched set,
racial categories, grade, ESL status, economic disadvantage, and sex,
along with a natural spline with five degrees of freedom for pretest.
This regression adjusts for imbalances in these covariates within
matched sets \citep[c.f.][]{biasAdjust}.
All three sets of standard errors
and confidence intervals used the heteroskedasticity-robust ``HC3'' sandwich estimator,
\citealt{sandwichPackage}.

The three estimates all roughly agree that being a high hint user
decreases posttest scores by roughly a fifth of a standard deviation.
That is, requesting hints hurts posttest scores---teachers should
recommend against students' requesting hints.

The similarity of the three estimates boosts our confidence, if only
slightly. The ``ATE'' and ``TOT'' estimates refer to different groups
of students---ATE estimates an average effect for the entire treatment
group, whereas TOT estimates an effect for high hint users. Had these
been very different, we might wonder why the effect of hint requests
varies so much between high and low hint users, and if that suggests
selection bias.
The fact that the additional covariate adjustment of the ANCOVA model
made little difference in the effect estimate suggests that the
propensity score match eliminated all or most of the confounding bias due to
measured covariates; hence, additional adjustment was unnecessary.

% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Dec 17 17:28:44 2019
\begin{table}[ht]
\centering
\begin{tabular}{rllccc}
  \hline
Weights & Estimate & Std. Error & CI & SI1 & SI2 \\ 
  \hline
OLS & -0.18 & 0.04 & [-0.25,-0.10] & [-0.41,0.06] & [-0.31,-0.04] \\ 
  ATE & -0.12 & 0.05 & [-0.23,-0.02] & [-0.45,0.20] & [-0.31,0.06] \\ 
  TOT & -0.14 & 0.06 & [-0.25,-0.03] & [-0.47,0.19] & [-0.33,0.05] \\ 
   \hline
\end{tabular}
\caption{Estimates with and without regression adjustment and sensitivity analysis of the effect of hint usage.} 
\label{tab:matchResults}
\end{table}

@@

None of these estimates addressed the possiblity of confounding from
unmeasured covariates; instead, they relied on matched ignorability.
This would be a particular concern if pretest scores do not completely
capture students' mathematics ability, and if weaker students ask for hints at
higher rates, and also score lower on the posttest.
Then, whatever aspects of student ability are not captures by the
pretest, and correlate with both hint requests and posttest scores,
would confound the estimates in the upper part of Table
\ref{tab:matchResults}.

To account for the possibility of unmeasured confounding, we conducted
a sensitivity analysis following the method of \citet{hhh}.
This method imagines a hypothetical missing covariate $U$ and estimates
how the the omission of $U$ alters the estimate
and its standard error.
The result is a ``sensitivity interval'' \citep[c.f.][]{rosenbaum2002observational} that,
with 95\% confidence, contains the true effect accounting for both
sampling uncertainty and uncertainty due to possible confounding.
$U$ is characterized by two sensitivity parameters:
$\rho^2$ is the squared partial correlation between $U$ and $Y$, after
accounting for the observed covariates in the model, and $T_Z$ is the
t-statistic from an ordinary least squares regression of $H$ on $U$, along with the observed covariates.
These sensitivity parameters may be benchmarked by estimating their
counterparts when observed covariates are left out of the model.
The first three rows of the table have both $\rho^2$ and $T_Z$ equal
to zero, to reflect that they assume no unmeasured confounders.
The row of Table \ref{tab:matchResults} labeled ``[Pretest]''
imagines a $U$ that predicts $Y$ and $H$ to a similar extent as
pretest.
The resulting sensitivity interval is quite wide, including
both large negative effects as well as substantial positive
effects.
A slightly less conservative sensitivity analysis, labeled
``[Pretest/2]'' allows $U$ to
predict $H$ and $\rho$ half as well as pretest scores.
This sensitivity interval is wide, but entirely negative.

@@

$\hat{\Delta}$ estimates the difference between the average outcome
for students in the entire treatment group, and the average outcome
for the treatment group were $H=0$ for all students.
That is
\begin{equation*}
\Delta=\EE[Y(1,H(1))-Y(1,0)|Z=1]
\end{equation*}
This can be re-written:
\begin{align*}
\begin{split}
\Delta&=\EE[Y(1,1)-Y(1,0)|Z=1,H=1]Pr(H=1)\\
&\qquad +\EE[Y(1,0)-Y(1,0)|Z=1,H=0]Pr(H=0)
\end{split}
\\[2ex]
&=\EE[Y(1,1)-Y(1,0)|Z=1,H=1]Pr(H=1)
&=Pr(H=1)TOT
\end{align*}
Since $Pr(H=1)=$0.7,
$\hat{\Delta}=$0.7$TOT$.
Estimated thusly, the indirect effect is negative: -0.1, with
a standard error of 0.04.
This analysis suggests that the availability of hints
actually \emph{lowers} the treatment effect, that CTA1 works not
because of hints, but despite them.

@@

We incorporated the model for hint requests, (\ref{eq:rasch}), into
the larger principal stratification model, replacing $\eta(1)$ for
$\bar{h}(1)$.
Then, the principal stratification estimand is
\begin{equation*}
\tau(r)=\EE[Y(1)-Y(0)|\eta(1)=r]
\end{equation*}
A full treatment of this variant of principal stratification, including a
discussion of identification and estimation, may be found in
\citet{aoas}.

Our approach to PS estimation is model-based and Bayesian, following
\citet{page2012principal}, \citet{feller2016compared},
\citet{schwartz2011bayesian} and others.
The PS model consists of three sub-models, all depending on a
 vector of parameters $\bm{\theta}$ that contains regression
 coefficients, variance components, and treatment effects.
The sub-models are:
$p(h;\eta(1),\bm{\theta})$, giving the distribution of actual hint
requests as a function of $\eta(1)$, $p(\eta(1)|\bm{x},\bm{\theta})$
giving the distribution of $\eta(1)$ conditional on covariates, and
$p(Y|\eta(1),Z,\bm{x},\bm{\theta})$ giving the conditional
distribution of outcomes.
With these three models in hand, and a prior distribution
$p(\bm{\theta})$, posterior inference for $\bm{\theta}$ proceeds based
on the following structure:
\begin{align*}
p(\bm{\theta}|\bm{Y},\bm{Z},&\bm{X},\{\bm{h}_i\}_{i:Z_i=1})\propto\\
p(\bm{\theta})&\displaystyle\prod_{i: Z_i=1}\int p(Y|\bm{x}_i,Z=1,\bm{\theta},\eta(1))p(\bm{h}_{i}|\eta(1)(1),\bm{\theta})p(\eta(1)|\bm{x}_i,\bm{\theta})d\eta(1)\\
\times&\displaystyle\prod_{i:Z_i=0} \int p(\yci|\bm{x}_i,\bm{\theta},\eta(1))p(\eta(1)|\bm{x}_i,\bm{\theta})d\eta(1).\numberthis\label{eq:posteriorEta}
\end{align*}
where $\bm{h}_i$ is the vector of hint request data for subject $i$,
ranging over all challenging problems.
In other words, we estimate parameters by integrating over unknown
(latent) $\eta(1)$ values, using the distribution of $\eta$ estimated
in the treatment group.
This is, in essence, an infinite mixture distribution, with outcome
distribution $p(Y|\cdot)$ and mixing proportions $p(\eta(1)|\cdot)$.
Unlike in typical PS setups, $\eta(1)$ is unobserved for both $Z=1$
and $Z=0$ treatment groups, but is estimated using both $\bm{h}$ and
$\bm{x}$ in the treatment group, but only $\bm{x}$ in the control
group.

@@

First, students' proclivity to request hints is measured by the $\eta$
parameter in (\ref{eq:rasch}).
In the next level, $\eta(1)$ is modeled as a function of baseline
covariates $\bm{x}$:
\begin{equation}\label{eq:rasch2}
\eta(1)|\left(\bm{x}_i,\bm{\theta}\right) \sim
\mathcal{N}\left(\bm{x_i}\bm{\beta}^M+\epsilon^{Mt}_{t[i]}+\epsilon^{Ms}_{s[i]},
\sigma^M\right)
\end{equation}
where $\bm{\beta}^U$ is a vector of coefficients.
Since students were nested within teachers, who were nested within
schools, we included normally-distributed school ($\epsilon_{Us}$) and
teacher ($\epsilon_{Ut}$) random intercepts.
The covariates in the model, $\bm{x}_i$, were detailed in Table
\ref{tab:covariateBalance}; preliminary model checking suggested
including a quadratic term for pretest, which was added as a column of
$\bm{x}_i$.

We modeled students' posttest scores $Y$ as
conditionally normal:
\begin{equation}\label{eq:outcomeSubmodel}
 Y|\left(Z_i,\bm{x}_i,\bm{\theta},\eta(1)\right) \sim  \mathcal{N}\left(
\beta^Y_{0b[i]}+\bm{x}_i^T\bm{\beta}^Y+a\eta(1)+Z_i\tau(\eta(1))+\epsilon^{Yt}_{t[i]}+\epsilon^{Ys}_{s[i]},\sigma^Y_{Z[i]}\right)
\end{equation}
where $\beta^Y_{0b[i]}$ is a fixed effect for $i$'s randomization block, $\bm{\beta}^Y$ are the
covariate coefficients, and $\epsilon^{Yt}$, and
$\epsilon^{Ys}$ are normally-distributed teacher and school random
intercepts.
The residual variance $\sigma^Y$ varies with treatment assignment $Z$;
this captures measurement error in $Y$, treatment effect heterogeneity
that is not linearly related to
$\eta(1)$, and other between-student variation in $Y$ that is not predicted by
the mean model.

Finally, we modeled treatment effects
$\tau(\eta(1))$ as linear:
\begin{equation}\label{eq:tauModel}
\EE[Y_{T}-Y_{C}|\eta(1)]=\tau({\eta(1)})=b_0+b_1\eta(1)
\end{equation}
While more complex models for $\tau(\eta(1))$ are theoretically
possible (for instance, \citet{jin2008principal} uses a quadratic
model), in our experience non-linear effect models do not perform as
well in model checks as the the linear model (\ref{eq:tauModel}).

@@

Observational studies, and mediation analyses that build on them,
require ignorability assumptions such as Strong Ignorability or
Matched Ignorability.
These may be particularly problematic in our study, since students who
are more likely to request hints are also more likely to struggle with
the material, and hence score lower on the posttest.
We addressed this concern in three ways: first, we focused our attention
on the set of problems in the tutor in which students either requested
a hint or made an error, second, with propensity score matching, and
third with sensitivity analysis.
None of these approaches is bulletproof---they each require further
assumptions about how students request hints.
Eliminating problems in which students showed no signs of struggling
is likely to reduce the association between hint requests and
underlying academic ability, but not eliminate it.
For instance, stronger students may be better able to figure out a
difficult problem without access to hints or error feedback than
weaker students.
Matching depends both on the adequacy of the observed covariates to
capture baseline differences between $H=0$ and $H=1$ students, and the
sensitivity analyses we performed assumed that any un-measured
covariate would predict $H$ no better than pretest, or no better than
a covariate half as important as pretest.
These assumptions are impossible to verify from the data at hand;
instead, they require a keen understanding of student practice within
the tutor.

@@

In contrast, no ignorabilty assumptions are necessary for principal
stratification.
That said, an assumption similar to ignorability may play an important
role for interpretation.
Further, principal stratification estimators require a model for $Y(0)$
conditional on $\eta(1)$, which is unobserved.
On the other hand, model misspecification is, in principle,
detectable---an assumption of a particular family of models is not
``untestable.''
Still, the data will, in general, be unable to distinguish between certain
alternative models for $Y(0)$ conditional on $\eta(1)$, that may in
turn lead to qualitatively different conclusions.
In practice, analysts make the untestable assumption that $Y(0)$ conditional on $\eta(1)$ is
distributed similarly to $Y(1)$ conditional on $\eta(1)$.

@@

Our observational study found that hint requests have a negative effect: students who
requested hints more often than the 70th
percentile score lower on the posttest, after controlling for observed covariates---as
far as we can tell, the optimal strategy is to never request a hint.
However, this observed effect may be due to confounding.
Mediation analysis essentially interprets this result in terms of the
overall CTA1 treatment effect.
Specifically, our analysis showed that CTA1 affected students'
posttest scores in (at least) two opposite ways: by allowing hints, it
lowered their posttest scores, but its other mechanisms increased
their scores by an even greater amount.

Like in the observational study and in mediation analysis, the
principal stratification analysis found a negative correlation
between hint requesting and posttest scores.
Unlike those two methods, PS understands $\eta(1)$ as a
student-specific baseline characteristic---albeit one that only
manifests in the treatment group.
Students who request more hints need more help, and thus score lower
on the posttest.
On the other hand, the PS analysis suggests that it's these students
who also receive more help---perhaps due to CTA1 hints, or other
features of the curriculum---and hence benefit more from CTA1.

>>>>

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

@@

Table \ref{tab:matchResults} gives treatment effect estimates,
standard errors, and confidence intervals under these three weighting
schemes.
All three sets of standard errors
and confidence intervals used the heteroskedasticity-robust ``HC3'' sandwich estimator,
\citealt{sandwichPackage}.
The three estimates all roughly agree that being a high hint user
decreases posttest scores by roughly a fifth of a standard deviation.
That is, requesting hints hurts posttest scores---teachers should
recommend against students' requesting hints.

The similarity of the three estimates boosts our confidence, if only
slightly. The ``ATE'' and ``TOT'' estimates refer to different groups
of students---ATE estimates an average effect for the entire treatment
group, whereas TOT estimates an effect for high hint users. Had these
been very different, we might wonder why the effect of hint requests
varies so much between high and low hint users, and if that suggests
selection bias.
% The fact that the additional covariate adjustment of the ANCOVA model
% made little difference in the effect estimate suggests that the
% propensity score match eliminated all or most of the confounding bias due to
% measured covariates; hence, additional adjustment was unnecessary.

% latex table generated in R 3.6.2 by xtable 1.8-4 package
% Tue Jan  7 14:23:39 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllccc}
  \hline
  &&&&\multicolumn{2}{c}{Sensitivity Intervals}\\
 Weights&Estimate&Std. Error&CI&[Pretest]&[Ethnicity] \\
 \hline
OLS & -0.18 & 0.04 & [-0.25,-0.10] & [-0.41,0.06] & [-0.30,-0.05] \\ 
  ATE & -0.12 & 0.05 & [-0.23,-0.02] & [-0.45,0.20] & [-0.29,0.04] \\ 
  TOT & -0.14 & 0.06 & [-0.25,-0.03] & [-0.47,0.19] & [-0.31,0.03] \\ 
   \hline
\end{tabular}
\caption{Estimates and sensitivity analysis of the effect of hint usage.} 
\label{tab:matchResults}
\end{table}

@@

None of these estimates addressed the possibility of confounding from
unmeasured covariates; instead, they relied on matched ignorability.
This would be a particular concern if pretest scores do not completely
capture students' mathematics ability, and if weaker students ask for hints at
higher rates, and also score lower on the posttest.
Then, whatever aspects of student ability are not captures by the
pretest, and correlate with both hint requests and posttest scores,
would confound the estimates in the upper part of Table
\ref{tab:matchResults}.

To account for the possibility of unmeasured confounding, we conducted
a sensitivity analysis following the method of \citet{hhh}.
This method imagines a hypothetical missing covariate $U$ and estimates
how the the omission of $U$ alters the estimate
and its standard error.
The result is a ``sensitivity interval'' \citep[c.f.][]{rosenbaum2002observational} that,
with 95\% confidence, contains the true effect accounting for both
sampling uncertainty and uncertainty due to possible confounding.
$U$ is characterized by two sensitivity parameters:
$\rho^2$ is the squared partial correlation between $U$ and $Y$, after
accounting for the observed covariates in the model, and $T_Z$ is the
t-statistic from an ordinary least squares regression of $H$ on $U$, along with the observed covariates.
These sensitivity parameters may be benchmarked by estimating their
counterparts when observed covariates are left out of the model.

The last two columns of table \ref{tab:matchResults} give sensitivity intervals for each of the three weighted average treatment effects.
The column of Table \ref{tab:matchResults} labeled ``[Pretest]''
accounts for the possible omission of a covariate $U$ that, at most, predicts $Y$ and $H$ as well as do pretest scores.
The resulting sensitivity intervals are quite wide, including
both large negative effects as well as substantial positive
effects.
In fact, pretest is the most important of our observed covariates; it is the observed covariate whose omission would cause the most bias.
This is unsurprising, since pretests are generally considered to be the most important covariate to measure, and studies that include pretest measures often reproduce experimental estimates \citep[e.g.][]{cook2008three,cook2009bias}.
Therefore, a sensitivity analysis considering an omitted covariate as important as pretest scores may be too pessimistic. 
Instead,
the final column of Table \ref{tab:matchResults} gives sensitivity intervals for the omission of a covariate that, at most, predicts $Y$ and $H$ as well as ethnicity dummy variables, the second most important of our observed covariates.
These sensitivity intervals are also wide, but the interval corresponding
to OLS weights is entirely negative.

The conclusion is that the omission of a confounder as important as ethnicity indicators
cannot explain the estimated OLS-weighted ATE.
Allowing for the possible omission of such a covariate, we infer that requesting a
large number of hints hurts students' posttest scores.
The size of the effect may be as small as
5\% of a standard deviation
or as large as 30\% of a standard deviation.

@@

$\hat{\Delta}$ estimates the difference between the average outcome
for students in the entire treatment group, and the average outcome
for the treatment group were $H=0$ for all students.
That is
\begin{equation*}
\Delta=\EE[Y(1,H(1))-Y(1,0)|Z=1]
\end{equation*}
This can be re-written:
\begin{align*}
\begin{split}
\Delta&=\EE[Y(1,1)-Y(1,0)|Z=1,H=1]Pr(H=1)\\
&\qquad +\EE[Y(1,0)-Y(1,0)|Z=1,H=0]Pr(H=0)
\end{split}
\\[2ex]
&=\EE[Y(1,1)-Y(1,0)|Z=1,H=1]Pr(H=1)\\
&=Pr(H=1)TOT
\end{align*}
Since $Pr(H=1)=$0.3,
$\hat{\Delta}=$0.3$TOT$.
Estimated thusly, the indirect effect is negative: -0.04, with
a standard error of 0.02.
This analysis suggests that the availability of hints
actually \emph{lowers} the treatment effect, that CTA1 works not
because of hints, but despite them.

@@

We incorporated the model for hint requests, (\ref{eq:rasch}), into
the principal stratification model, replacing $\eta(1)$ for
$\bar{h}(1)$.
Then, the principal stratification estimand is
\begin{equation*}
\tau(r)=\EE[Y(1)-Y(0)|\eta(1)=r]
\end{equation*}
A full treatment of this variant of principal stratification, including a
discussion of identification and estimation, may be found in
\citet{aoas}.

Our approach to PS estimation is model-based and Bayesian, following
\citet{page2012principal}, \citet{feller2016compared},
\citet{schwartz2011bayesian} and others.
The PS model consists of three sub-models, all depending on a
 vector of parameters $\bm{\theta}$ that contains regression
 coefficients, variance components, and treatment effects.
The sub-models are:
$p(h;\eta(1),\bm{\theta})$, giving the distribution of actual hint
requests as a function of $\eta(1)$, $p(\eta(1)|\bm{x},\bm{\theta})$
giving the distribution of $\eta(1)$ conditional on covariates, and
$p(Y|\eta(1),Z,\bm{x},\bm{\theta})$ giving the conditional
distribution of outcomes.
With these three models in hand, and a prior distribution
$p(\bm{\theta})$, posterior inference for $\bm{\theta}$ proceeds based
on the following structure:
\begin{align*}
p(\bm{\theta}|\bm{Y},\bm{Z},&\bm{X},\{\bm{h}_i\}_{i:Z_i=1})\propto\\
p(\bm{\theta})&\displaystyle\prod_{i: Z_i=1}\int p(Y_i|\bm{x}_i,Z=1,\bm{\theta},\eta(1))p(\bm{h}_{i}|\eta(1),\bm{\theta})p(\eta(1)|\bm{x}_i,\bm{\theta})d\eta(1)\\
\times&\displaystyle\prod_{i:Z_i=0} \int p(Y_i|\bm{x}_i,Z=0,\bm{\theta},\eta(1))p(\eta(1)|\bm{x}_i,\bm{\theta})d\eta(1).\numberthis\label{eq:posteriorEta}
\end{align*}
where $\bm{h}_i$ is the vector of hint request data for subject $i$,
ranging over all challenging problems.
In other words, we estimate parameters by integrating over unknown
(latent) $\eta(1)$ values, using the distribution of $\eta$ estimated
in the treatment group.
This is, in essence, an infinite mixture distribution, with outcome
distribution $p(Y|\cdot)$ and mixing proportions $p(\eta(1)|\cdot)$.
Unlike in typical PS setups, $\eta(1)$ is unobserved for both $Z=1$
and $Z=0$ treatment groups, but is estimated using both $\bm{h}$ and
$\bm{x}$ in the treatment group, but only $\bm{x}$ in the control
group.

@@

First, students' proclivity to request hints is measured by the $\eta$
parameter in (\ref{eq:rasch}).
In the next level, $\eta(1)$ is modeled as a function of baseline
covariates $\bm{x}$:
\begin{equation}\label{eq:rasch2}
\eta(1)|\left(\bm{x}_i,\bm{\theta}\right) \sim
\mathcal{N}\left(\bm{x_i}\bm{\beta}+\epsilon^{tch}_{t[i]}+\epsilon^{scl}_{s[i]},
\sigma\right)
\end{equation}
where $\bm{\beta}$ is a vector of coefficients.
Since students were nested within teachers, who were nested within
schools, we included normally-distributed school ($\epsilon^{scl}$) and
teacher ($\epsilon^{tch}$) random intercepts.
The covariates in the model, $\bm{x}_i$, were detailed in Table
\ref{tab:covariateBalance}; preliminary model checking suggested
including a quadratic term for pretest, which was added as a column of
$\bm{x}_i$.

We modeled students' posttest scores $Y$ as
conditionally normal:
\begin{equation}\label{eq:outcomeSubmodel}
 Y|\left(Z_i,\bm{x}_i,\bm{\theta},\eta(1)\right) \sim  \mathcal{N}\left(
\gamma_{0b[i]}+\bm{x}_i^T\bm{\gamma}+a\eta(1)+Z_i\tau(\eta(1))+\zeta^{tch}_{t[i]}+\zeta^{scl}_{s[i]},\omega_{Z[i]}\right)
\end{equation}
where $\bamma_{0b[i]}$ is a fixed effect for $i$'s randomization block, $\bm{\gamma}$ are the
covariate coefficients, and $\zeta^{tch}$, and
$\zeta^{scl}$ are normally-distributed teacher and school random
intercepts.
The residual variance $\omega$ varies with treatment assignment $Z$;
this captures measurement error in $Y$, treatment effect heterogeneity
that is not linearly related to
$\eta(1)$, and other between-student variation in $Y$ that is not predicted by
the mean model.

Finally, we modeled treatment effects
$\tau(\eta(1))$ as linear:
\begin{equation}\label{eq:tauModel}
\EE[Y(1)-Y(0)|\eta(1)]=\tau({\eta(1)})=b_0+b_1\eta(1)
\end{equation}
While more complex models for $\tau(\eta(1))$ are theoretically
possible (for instance, \citet{jin2008principal} uses a quadratic
model), in our experience non-linear effect models do not perform as
well in model checks as the the linear model (\ref{eq:tauModel}).

@@

Observational studies, and mediation analyses that build on them,
require ignorability assumptions such as Strong Ignorability or
Matched Ignorability.
These may be particularly problematic in our study, since students who
are more likely to request hints are also more likely to struggle with
the material, and hence score lower on the posttest.
We addressed this concern in three ways: first, we focused our attention
on the set of problems in the tutor in which students either requested
a hint or made an error, second, with propensity score matching, and
third with sensitivity analysis.
None of these approaches is bulletproof---they each require further
assumptions about how students request hints.
Eliminating problems in which students showed no signs of struggling
is likely to reduce the association between hint requests and
underlying academic ability, but not eliminate it.
For instance, stronger students may be better able to figure out a
difficult problem without access to hints or error feedback than
weaker students.
Matching depends both on the adequacy of the observed covariates to
capture baseline differences between $H=0$ and $H=1$ students, and the
sensitivity analyses we performed assumed that any un-measured
covariate would predict $H$ and $Y$ no better than pretest or ethnicity indicators, respectively.
These assumptions are impossible to verify from the data at hand;
instead, judging their plausibility requires a keen understanding of student practice within
the tutor.

@@

In contrast, no ignorabilty assumptions are necessary for principal
stratification.
% That said, an assumption similar to ignorability may play an important
% role for interpretation.
However, principal stratification estimators require modeling
assumptions---most problematically, $p(Y|\bm{x},Z=0,\bm{\theta},\eta(1))$ of
\eqref{eq:posteriorEta}, the model for $Y(0)$
conditional on $\eta(1)$, $\bm{x}$, and parameters $\bm{\theta}$.
While any model specification can, and should, be checked against
observed data, this process can only narrow space of acceptable
models, it cannot be used to determine the correct form of
$p(Y|\bm{x},Z=0,\bm{\theta},\eta(1))$.
This is because $Y(0)$ and $\eta(1)$ are never observed simultaneously.
Therefore, it is possible for an analyst to misspecify $p(Y|\bm{x},Z=0,\bm{\theta},\eta(1))$, such that principal effect estimates are
severely biased, without being able to detect the model
misspecification with the data.
That is, the data will typically be unable to distinguish between
alternative $p(Y|\bm{x},Z=0,\bm{\theta},\eta(1))$ models that lead to
qualitatively different conclusions.
In practice, analysts make the untestable assumption that $p(Y|\bm{x},Z=0,\bm{\theta},\eta(1))$ is
of the same form as $p(Y|\bm{x},Z=1,\bm{\theta},\eta(1))$.

@@

Our observational study suggested that hint requests have a negative effect: students who
requested hints more often than the 70th
percentile score lower on the posttest, after controlling for observed
covariates.
As far as we can tell, the optimal strategy is to request few, if any, hints.
While our estimate is subject to bias from an omitted confounder, the omitted variable must be a relatively strong predictor of hint usage and of posttest scores in order to explain the observed relationship.
An omitted confounder that is as important as pretest could explain the relationship, but an omitted confounder as important as ethnicity---our second most important covariate---could not (so long as we estimate an OLS-weighted average effect).

Mediation analysis essentially interprets the observational study result in terms of the
overall CTA1 treatment effect.
Specifically, our analysis showed that CTA1 affected students'
posttest scores in (at least) two opposite ways: by allowing hints, it
lowered their posttest scores, but its other mechanisms increased
their scores by an even greater amount.

Like in the observational study and in mediation analysis, the
principal stratification analysis found a negative correlation
between hint requesting and posttest scores.
Students who request more hints need more help, and thus score lower
on the posttest.
Fortunately, the PS analysis also suggests that these students
receive more help----that is, they benefit more from CTA1---than do their peers who request fewer hints.
However, it is unclear whether their larger treatment effects are due to their higher rate of hint requests, or to some other correlated factor.
